---
title: "Biofilm Traits"
author: "Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

A trait based approach to bacterial biofilms in soils using *Pseudomonas aeruginosa* as a model system

```{r}
rm(list=ls())
getwd()
setwd("~/GitHub/BiofilmTrait")
```

Load package for writing figures
```{r}
require("png")
```

## SUPPL. FIG. 1: BIOFILM VS. LAG

Read in trait data from Lennon et al. (2012)
```{r}
setwd("~/GitHub/BiofilmTrait")
traits <- read.table("data/Biofilm_Lennon_2012.txt", sep="\t", header=TRUE)
```

Remove rows with NAs
```{r}
traits <- traits[complete.cases(traits),]
```

Plot biofilm and lag time
```{r}
png(filename="~/GitHub/BiofilmTrait/figures/SupplementalFigure1.png",
    width = 1200, height = 1200, res = 96*2)

par(mar = c(5, 6, 4, 2))
plot(traits$lag_time,traits$biofilm, xlab = "Lag time (hr)",
     ylab = expression('Biofilm Production (a'[550]*')'), pch = 22, 
     cex = 3, col = "black", cex.lab = 2, las = 1, lwd = 2,
     yaxt = "n", xaxt = "n")
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c("0.0", "1.0", "2.0", "3.0"), at = c(0, 1.0, 2.0, 3.0))
 
axis(side = 4, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    at = c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0, 3.5), labels = F)
 
axis(side = 1, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c("0", "20", "40", "60", "80"), at = c(0, 20, 40, 60, 80))

axis(side = 3, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    at = c(0, 20, 40, 60, 80), labels = F)

# Add p-value for t-test
mtext(expression(~italic(rho)~"= 0.38"), line = -2, cex = 1, at = 70)
mtext(expression(~italic("P")~"= 0.015"), line = -3, cex = 1, at = 70)

dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
```

Install package Hmisc and perform correlation
```{r}
library(Hmisc)
rcorr(traits$lag_time,traits$biofilm, type = "pearson")
```

The relationship is only marginally significant (r = -0.27, P = 0.098).
However, the relationship doesn't look linear. 
Use a Spearman rank correlation (non-parametric): r = -0.38, P = 0.0149. 

```{r}
rcorr(traits$lag_time,traits$biofilm, type = "spearman")
```

## GROWTH RATES UNDER LAB CONDITIONS

Read OD600 data from cultures to cacluate Malthusian growth rates
```{r}
setwd("~/GitHub/BiofilmTrait")
malth <- read.csv("data/Biofilm_Malthusian.csv",sep = ",", header=TRUE)
```

Define Time Points
```{r}
t0 <- malth[which(malth$Time_h==0),]
t38 <- malth[which(malth$Time_h==38),]
data <- cbind(t0,t38)

data_growth <- data[ -c(1,6,8,9) ]
colnames(data_growth)[1] <- "t0"
colnames(data_growth)[4] <- "OD_t0"
colnames(data_growth)[5] <- "t38"
colnames(data_growth)[6] <- "OD_t38"
```

Calculate growth rates and conduct t-test
```{r}
rate <- (log(data_growth$OD_t38) - log(data_growth$OD_t0)) / 
        (data_growth$t38-data_growth$t0)
growth <- cbind(data_growth,rate)
malth.t.test <- t.test(growth$rate ~ growth$Strain)
```

No effect of strain on growth rate under laboratory maintenance conditions. 
Results qualitatively independent of sampling interval. 
Welch Two-Sample t-test: t = -1.118, df = 2.217, p-value = 0.37

## FIG. 1: MULTI-PANEL BIOFILM 

Panel A: Comparison of 'NM' and 'OE' with O'Toole assay 

Load data, calculate group means and SEM, run t-test
```{r}
setwd("~/GitHub/BiofilmTrait")
otoole <- read.csv("data/Biofilm_Otoole.csv", head = TRUE, sep = ",")
sem <- function(x) sqrt(var(x)/length(x))
strain <- factor(otoole$Strain, levels = c('non-mucoid', 'mucoid'))
otoole.means <- tapply(otoole$Biofilm, strain, mean)
otoole.sem <- tapply(otoole$Biofilm, strain, sem)
otoole.t.test <- t.test(otoole$Biofilm ~ otoole$Strain)
```

Plot biofilm data
```{r}
# Set-up for printing 
png(filename="~/GitHub/BiofilmTrait/figures/Figure1.png",
    width = 800, height = 1200, res = 96*2)

# Set plotting parameters
par(mar = c(1, 8, 4, 8))
layout <- layout(rbind(1, 2, 3), height = c(3, 0.8, 3)) 
#layout.show(layout)

# Make inital plot
otoole.plot <- plot(otoole.means, log = "y", ylim = c(0.05, 2.4), 
      xlim = c(0.5, 2.5), pch = 22, bg = c("white", "grey"), lwd = 2, 
      cex = 3, yaxt = "n", xaxt = "n", cex.lab = 2, cex.axis = 1.5, 
      las = 1, ylab = "", xlab = "")
      
      box(lwd = 2)

# Add y-label axis for NM vs. OE plot
mtext(expression('Biofilm Production (a'[550]*')'), side = 2, 
      outer = TRUE, cex = 1, line = -4, adj = 0.87)

# Add p-value for t-test
mtext(expression(~italic("P")~"< 0.001"), line = -1.75, cex = 1, at = 0.9)
text(2.4, 2, labels = "A", cex = 2)

# Major Axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c(0.1, 1), at = c(0.1, 1))
 
axis(side = 4, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    at=c(0.1, 1), labels = F)
 
axis(side = 1, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c("NM", "OE"), at = c(1, 2))

axis(side = 3, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    at = c(1, 2), labels = F)

# Minor Axes
axis(side = 2, tck = -0.015, labels = F, lwd.ticks = 2,
       at = c(0.05, 0.1, 0.25, 0.5, 2))

axis(side = 4, tck = -0.015, labels = F, lwd.ticks = 2,
       at = c(0.05, 0.1, 0.25, 0.5, 2))

# Load colony images
setwd("~/GitHub/BiofilmTrait")
NM.img <- readPNG("./data/NM.png")
grid.raster(NM.img, x = 0.38, y = 0.48, height = 0.1)

OE.img <- readPNG("./data/OE.png")
grid.raster(OE.img, x = 0.61, y = 0.48, height = 0.1)

# Load biofilm density curve
setwd("~/GitHub/BiofilmTrait")
kern.data <-read.csv(file = "./data/Biofilm_Spp.csv", head=TRUE, sep =",")
kern <- density(kern.data$Biofilm) 

plot.new()
par(mar = c(4, 8, 2, 8))
plot(kern, main = NA, xaxt = "n", yaxt = "n", cex.lab = 1.5, ylab = "",
     xlab = "", xlim = c(-0.75, 3.8), ylim = c(0,1), lwd = 2)

mtext('Density', side = 2, outer = TRUE, cex = 1, 
      line = -4, adj = 0.25)

mtext(expression('Biofilm Production (a'[550]*')'), side = 1, outer = TRUE, 
      cex = 1, line = -1, adj = 0.5)
    
     axis(side = 1, lwd.ticks = 2, cex.axis = 1.5, las = 1,
        at = c(0, 1, 2, 3), labels = T)

     axis(side = 3, lwd.ticks = 2, cex.axis = 1.5, las = 1,
        at = c(0, 1, 2, 3), labels = F)

     axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
        at = c(0, 0.5, 1.0), labels = T)

    axis(side = 4, lwd.ticks = 2, cex.axis = 1.5, las = 1,
        at = c(0, 0.5, 1.0), labels = F)

box(lwd=2)

arrows(0.2, 0.6, 0.2, 0.8, length = 0.05, col = "Black")
text(0.2, 0.55, labels =" NM", cex = 0.75)
arrows(3, 0.32, 3, 0.12, length = 0.05, col = "Black") 
text(3, 0.36, labels =" OE", cex = 0.75)
text(3.6, 0.95, labels = "B", cex = 2)

dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
```

## FIG. 2: SURVIVORSHIP 
```{r}
require(survival)
require(extrafont)
setwd("~/GitHub/BiofilmTrait")
surv <- read.table("./data/Biofilm_Survivorship.txt",header=TRUE,sep="\t")
biofsurv <- Surv(surv$Days,surv$Event)
biofsurv.fit <- survfit(biofsurv ~ surv$Group, conf.int = TRUE, 
                type = "kaplan-meier")
biofsurv.diff <- survdiff(Surv(surv$Days,surv$Event) ~ surv$Group, data = surv)
```

Plotting
```{r}
png(filename="~/GitHub/BiofilmTrait/figures/Figure2.png",
    width = 1200, height = 1200, res = 96*2)

par(mar = c(5, 7, 5, 7))
plot(biofsurv.fit, conf.int = TRUE, mark.time = FALSE,
  xlim = c(0,50), ylim = c(0,1),
  lty = c(1,3,3,1,3,3),
  col = c("black", "grey", "grey", "black", "grey", "grey"),
  xlab = "Time (d)",
  ylab = "", cex.lab = 1.5, cex.axis = 1.2, las = 1, lwd = 2,
  yaxt = "n", xaxt = "n")
  box(lwd=2)

mtext("Survivorship", side = 2, outer = TRUE, cex = 1.5, line = -3, adj = 0.5)

axis(side = 2, labels = T, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
       at = c(0, 0.25, 0.5, 0.75, 1.0))
axis(side = 4, labels = F, lwd.ticks = 2, 
       at = c(0, 0.25, 0.5, 0.75, 1.0))
axis(side = 1, labels = T, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
       at = c(0, 10, 20, 30, 40, 50))
axis(side = 3, labels = F, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
       at = c(0, 10, 20, 30, 40, 50))

text(44, 0.65, "OE", cex = 1.25)
text(25, 0.41, "NM", cex = 1.25)

dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
```

## FIG. 3: GROWTH VS. WATER POTENTIAL

Load data, peform multiple regressions, and calculate Psi stars
```{r}
# Load data
setwd("/Users/lennonj/GitHub/BiofilmTrait")
growth <- read.table("./data/Biofilm_Growth.txt",header=TRUE,sep="\t")

# Specify variables
MPa <- growth[,1]
Growth <- growth[,2]
Strain <- growth[,3]

# Create Dummy Variables by Strain
D1 <- (Strain == "NM")*1
fit <- lm(Growth ~ MPa + D1 + MPa*D1)
summary(fit)

# Multiple regression parameters
OE.int <- fit$coefficients[1]
OE.slp <- fit$coefficients[2]
NM.int <- OE.int + fit$coefficients[3]
NM.slp <- OE.slp + fit$coefficients[4]

# Strain-specific values
NM <- growth[which(growth$Strain == "NM"),]
OE <- growth[which(growth$Strain == "OE"),]

# Plotting
png(filename="/Users/lennonj/GitHub/BiofilmTrait/figures/Figure3.png",
    width = 1200, height = 1200, res = 96*2)
par(mar = c(7, 7, 5, 7))
plot(jitter(NM$MPa, factor = 10), NM$Growth, xlim = c(max(growth$MPa + 0.1), 
  min(growth$MPa - 0.2)), ylim = c(-3.5, 3),
  pch = 22, bg = "white", col = "black", cex = 2,
  ylab = "", xlab = "", cex.lab = 1.5, cex.axis = 1.2, 
  las = 1, lwd = 2, yaxt = "n", xaxt = "n")
  box(lwd=2)

# Add points
points(jitter(OE$MPa, factor = 10), OE$Growth, pch = 22, bg = "grey", 
       col = "black", cex = 2, lwd = 2)

# Add axis labels
mtext(expression('Growth Rate (d'^-1*')'), side = 2, outer = TRUE, 
      cex = 1.5, line = -3, adj = 0.5)
mtext("Water Potential (MPa)", side = 1, outer = TRUE, cex = 1.5, 
      line = -3, adj = 0.5)

# Add ticks and tick labels
axis(side = 2, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   labels = c("-3.0", "-1.5", "0.0", "1.5", "3.0"), at = c(-3, -1.5, 0.0, 1.5, 3.0))

axis(side = 4, labels = F, lwd.ticks = 2, 
   at = c(-3, -1.5, 0.0, 1.5, 3.0))

axis(side = 1, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   labels = c("0.0", "-0.5", "-1.0", "-1.5"), at = c(0, -0.5, -1, -1.5))

axis(side = 3, labels = F, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   at = c(0, -0.5, -1, -1.5))

# Add Psi star lines 
NM.star <- -NM.int/NM.slp
segments(x0 = NM.star, y0 = -3.3, x1 = NM.star, y1 = 0, col = "black", 
         lwd = 2.5, lty = 6, xpd = T)

OE.star <- -OE.int/OE.slp
segments(x0 = OE.star, y0 = -3.3, x1 = OE.star, y1 = 0, col = "black", 
         lwd = 2.5, lty = 4, xpd = T)

# Add zero-growth line 
abline(h = 0, col = "black", lty = 3, lwd = 2)

# Add multiple regression lines
clip(0.05,-1.6,-3.5,3)
abline(a = NM.int, b = NM.slp, col = "black", lwd = 2.5, lty = 6)

clip(0.05,-1.6,-3.5,3)
abline(a = OE.int, b = OE.slp, col = "black", lwd = 2.5, lty = 4)

# Add legend
legend(-1.2, 3.24, c("OE", "NM"), pch = 22, pt.bg = c("grey", "white"), 
       pt.cex = 2, pt.lwd = 2, bty = 'n', y.intersp = 1, lty = c(6,4), 
       lwd = 2.5)

# Add Psi star symbols to plot
mtext(side = 1, line = -1, at = NM.star, bquote(psi~'*'[NM]), cex = 1.2)
mtext(side = 1, line = -1, at = OE.star, bquote(psi~'*'[OE]), cex = 1.2)

# Add R2
text(-0.075, 2.9, labels = expression(R^2 == 0.65), cex = 1)

dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
```

## FIG. 4: RESPIRATION VS. WATER POTENTIAL

Load data and run MLE analyses for multiple model comparisons

```{r}
setwd("~/GitHub/BiofilmTrait")
resp.raw <- read.csv("./data/Biofilm_Respiration.csv", header = TRUE, sep =",")
resp.trunc <- resp.raw[,1:3] # gets rid of trailing data
resp.uneg <- resp.raw[,1]*-1 # convert MPa to positive values
resp <- data.frame(resp.uneg,resp.trunc) # add uneg to dataframe
colnames(resp)[1] <- "Wp" # "Wp" = uneg MPa (positive)
require(bbmle)

#starting values for niche model
A = 35 # Maximum respiratoin
X = 0 # Optimum MPa
B = 0.05 # Niche Breadth
T = 1 # Tau, shape kernel
Z = 7.5 # Error

fit1 <- mle2(R ~ dnorm(mean = a, sd = z), start = list(a = A, z = Z),
             data = resp) 
             # null

fit2 <- mle2(R ~ dnorm(mean = a * exp(-((abs(W + x))/b)^t), sd = z),
             method = "L-BFGS-B", lower = c(x = 0.0000, b = 0.001), 
             start = list(a = A, x = X, b = B, t = T, z = Z), data = resp) 
             # common parameter set for both strains

fit3 <- mle2(R ~ dnorm(mean = a * exp(-((abs(W + x))/b)^t), sd = z),
             method = "L-BFGS-B", lower = c(x = 0.0000, b = 0.001), 
             start = list(a = A, x = X, b = B, t = T, z = Z), data = resp, 
             parameters = c(a ~ Strain)) 
             # max differs between strains

fit4 <- mle2(R ~ dnorm(mean = a*exp(-((abs(W + x))/b)^t), sd = z), 
             method = "L-BFGS-B",lower = c(x = 0.0000, b = 0.001), 
             start = list(a = A, x = X, b = B,t = T, z = Z), 
             data = resp, parameters = c(x ~ Strain)) 
             # opt water differs between strain 

fit5 <- mle2(R ~ dnorm(mean = a*exp(-((abs(W + x))/b)^t), sd = z), 
             method = "L-BFGS-B", lower = c(x = 0.0000, b = 0.001), 
             start = list(a = A, x = X, b = B, t = T, z = Z), 
             data = resp, parameters = c(b ~ Strain)) 
             # breadth differs between strains

fit6 <- mle2(R ~ dnorm(mean = a*exp(-((abs(W + x))/b)^t), sd = z), 
             method = "L-BFGS-B", lower = c(x = 0.0000, b = 0.001), 
             start = list(a = A, x = X, b = B, t = T, z = Z), 
             data = resp, parameters = c(t ~ Strain)) 
              # tau differs between strains

fit7 <- mle2(R ~ dnorm(mean = a*exp(-((abs(W + x))/b)^t), sd = z), 
             method = "L-BFGS-B",lower = c(x = 0.0000, b = 0.001), 
             start = list(a = A, x = X, b = B, t = T, z = Z), 
             data = resp, parameters = c(a ~ Strain, x ~ Strain)) 
             # max and opt differ between strains

fit8 <- mle2(R ~ dnorm(mean = a*exp(-((abs(W + x))/b)^t), sd = z), 
             method = "L-BFGS-B", lower = c(x = 0.0000, b = 0.001), 
             start = list(a = A, x = X, b = B, t = T, z = Z), 
             data = resp, parameters = c(a ~ Strain, b ~ Strain)) 
             # max and breadth differ between strains

fit9 <- mle2(R ~ dnorm(mean = a*exp(-((abs(W + x))/b)^t), sd = z), 
             method = "L-BFGS-B", lower = c(x = 0.0000, b = 0.001), 
             start = list(a = A, x = X, b = B, t = T, z = Z), 
             data = resp, parameters = c(a ~ Strain, t ~ Strain)) 
             # max and tau differ between strains

fit10 <- mle2(R ~ dnorm(mean = a*exp(-((abs(W + x))/b)^t), sd = z), 
              method = "L-BFGS-B", lower = c(x = 0.0000, b = 0.001), 
              start = list(a = A, x = X, b = B, t = T, z = Z), 
              data = resp, parameters = c(x ~ Strain, b ~ Strain)) 
              # opt and breadth differ between strains

fit11 <- mle2(R ~ dnorm(mean = a*exp(-((abs(W + x))/b)^t), sd = z), 
              method = "L-BFGS-B", lower = c(x = 0.0000, b = 0.001), 
              start = list(a = A, x = X, b = B, t = T, z = Z), 
              data = resp, parameters = c(x ~ Strain, t ~ Strain)) 
              # opt and tau differ between strains

fit12 <- mle2(R ~ dnorm(mean = a*exp(-((abs(W + x))/b)^t), sd = z), 
              method = "L-BFGS-B", lower = c(x = 0.0000, b = 0.001), 
              start = list(a = A, x = X, b = B, t = T, z = Z), 
              data = resp, parameters = c(b ~ Strain, t ~ Strain)) 
              # breadth and tau differ between strains

fit13 <- mle2(R ~ dnorm(mean = a*exp(-((abs(W + x))/b)^t), sd = z), 
              method = "L-BFGS-B", lower = c(x = 0.0000, b = 0.001), 
              start = list(a = A, x = X, b = B, t = T, z = Z), 
              data = resp, parameters = c(a ~ Strain, x ~ Strain, 
              b ~ Strain)) 
              # max, opt, breadth differ between strains

fit14 <- mle2(R ~ dnorm(mean = a*exp(-((abs(W + x))/b)^t), sd = z), 
              method = "L-BFGS-B", lower = c(x = 0.0000, b = 0.001), 
              start = list(a = A, x = X, b = B, t = T, z = Z), 
              data = resp, parameters = c(a ~ Strain, x ~ Strain, 
              t ~ Strain)) 
              # max, opt, tau differ between strains

fit15 <- mle2(R ~ dnorm(mean= a*exp(-((abs(W + x))/b)^t), sd = z), 
              method = "L-BFGS-B", lower = c(x = 0.0000, b = 0.001), 
              start = list(a = A, x = X, b = B, t = T, z = Z), 
              data = resp, parameters = c(a ~ Strain, b ~ Strain, 
              t ~ Strain)) 
              # max, breadth, tau differ between strains

fit16 <- mle2(R ~ dnorm(mean = a*exp(-((abs(W + x))/b)^t), sd = z), 
              method = "L-BFGS-B", lower = c(x = 0.0000, b = 0.001), 
              start = list(a = A, x = X, b = B, t = T,
              z = Z), data = resp, parameters = c(x ~ Strain, 
              b ~ Strain, t ~ Strain)) 
              # opt, breadth, tau differ between strains

fit17 <- mle2(R ~ dnorm(mean = a*exp(-((abs(W + x))/b)^t), sd = z), 
              method = "L-BFGS-B", lower = c(x = 0.0000, b = 0.001), 
              start = list(a = A, x = X, b = B, t = T, z = Z), 
              data = resp, parameters = c(a ~ Strain, x ~ Strain, 
              b ~ Strain, t ~ Strain)) 
              #all params differ between strains

AIC.table <- AICtab(fit1, fit2, fit3, fit4, fit5, fit6, fit7, fit8, fit9, fit10,
              fit11, fit12, fit13, fit14, fit15, fit16, fit17, delta = T, 
              weights = T)
AIC.names <- attributes(AIC.table)$row.names
AIC.table <- as.data.frame(do.call(cbind, c(AIC.table)))
AIC.table$names <- AIC.names
write.table(AIC.table, "~/GitHub/BiofilmTrait/figures/AIC.table.txt", 
            sep = "\t") 

# Conclusion based on AIC is that the following models all perform equally
# well (i.e., deltaAIC ~ < 2): fit10, fit 4, fit7, fit16, fit13, and fit11. 
# However, all of these models contain Wopt. Therefore, discuss fit4, which only
# distinguishes OE and NM based on Wopt. 

NM.opt <- fit4@coef[2]
OE.opt <- NM.opt + fit4@coef[3]
```

#Calcuate niche breadth
Based on Lennon et al. (2012), we estimate niche breadth (`nb`) using the `b` parameter from the equation above. 

```{r}
# Following calculates estimates and error for `Wopt`, optmimum water potential
NM.opt <- fit4@coef[2]
NM.opt.e <- coef(summary(fit4))[2,2] 
OE.opt <- fit4@coef[3] + fit4@coef[2]
OE.opt.e <- NM.opt.e + coef(summary(fit4))[3,2] 
```

# Plot Data
```{r}
OE.resp <- subset(resp[,2:3], resp[,4] == "OE", data = resp)
NM.resp <- subset(resp[,2:3], resp[,4] == "NM", data = resp)

png(filename="~/GitHub/BiofilmTrait/figures/Figure4.png",
    width = 1200, height = 1200, res = 96*2)

plot.new()
par(mar = c(7, 7, 5, 7))

plot(jitter(OE.resp[,1], factor = 10), OE.resp[,2], xlim = c(0.5, -3.5), 
     ylim = c(-2.5, 40), type = "p", 
     pch = 22, bg = "grey", col = "black", cex = 2, ylab = "", xlab = "", 
     cex.lab = 1.5, las = 1, lwd = 2, yaxt = "n", xaxt = "n")
box(lwd=2)

points(jitter(NM.resp[,1], factor = 10), NM.resp[,2], type = "p", pch = 22, 
       bg = "white", col = "black",
      cex = 2, cex.lab = 1.5, lwd = 2)

# Add ticks and tick labels
axis(side = 2, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   labels = c(0, 10, 20, 30, 40), at = c(0, 10, 20, 30, 40))

axis(side = 4, labels = F, lwd.ticks = 2, 
   at = c(0, 10, 20, 30, 40))

axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1, mgp = c(3, 1, 0),
    labels = c(0, -1, -2, -3, -4), at = c(0, -1, -2, -3, -4))

axis(side = 3, labels = F, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   at = c(0, -1, -2, -3, -4))

mtext('Water Potential (MPa)', side = 1, outer = TRUE, cex = 1.5, 
      line = -4, adj = 0.5)

mtext(expression(paste('Respiration (', mu ,'gC-CO'[2]* ' g soil'^-1* 'd'^-1*')')), 
      side = 2, outer = TRUE, cex = 1.5, line = -3.5, adj = 0.6)

legend(-2, 40, c("OE", "NM"), pch = 22, pt.bg = c("grey", "white"), pt.cex = 2, 
       pt.lwd = 2, bty = 'n', y.intersp = 1, lty = c(4,6), lwd = 2.5, seg.len = 5)

# Add functions to plot

curve(coef(fit4)[[1]]*exp(-((abs(coef(fit4)[[2]]+x)/coef(fit4)[[4]]))^coef(fit4)[[5]]), 
      from = 0.2, to = -3.3, add = TRUE, lty = 6, lwd = 2.5) # adds NM

# For second curve, need to modify to get OE strain parameters

OEx=coef(fit4)[2]+coef(fit4)[3] # modifies optimum parameter relative to NM

curve(coef(fit10)[[1]]*exp(-((abs(OEx+x)/coef(fit4)[[4]])^coef(fit4)[[5]])), 
      from = 0.2, to= -3.3, add = TRUE, lty = 4, lwd = 2.5) # adds OE

dev.off()
graphics.off()
```

## FIG. 5: REWETTING COLUMN EXPERIMENT

# Read and arrange data
```{r}
setwd("~/GitHub/BiofilmTrait")
cols <- read.csv("./data/Biofilm_Columns.csv", header = TRUE, sep =",")
head(cols)

sem <- function(x) sqrt(var(x)/length(x))

OE.cols <- data.frame(cols[,2], apply(cols[, 3:6], 1, mean), 
                      apply(cols[, 3:6], 1, sem))
colnames(OE.cols)[1:3] <- c("day", "OE.mean", "OE.sem")

NM.cols <- data.frame(cols[,2], apply(cols[, 7:10], 1, mean), 
                      apply(cols[, 7:10], 1, sem))
colnames(NM.cols)[1:3] <- c("day", "NM.mean", "NM.sem")
```

# Make main plot
```{r}
png(filename="~/GitHub/BiofilmTrait/figures/Figure5.png",
    width = 1200, height = 800, res = 96*2)

plot.new()
par(mar = c(7, 7, 5, 7))

plot(OE.cols[,1], OE.cols[,2], xlim = c(0,100), ylim = c(-0.017, 0.001), type = "b", 
     pch = 22, bg = "grey", col = "black", cex = 2, ylab = "", xlab = "", 
     cex.lab = 1.5, las = 1, lwd = 2, yaxt = "n", xaxt = "n",
     panel.first={
       arrows(x0 = OE.cols[,1], y0 = OE.cols[,2], y1 = OE.cols[, 2] - OE.cols[, 3],
              angle = 90, length = 0.05, lwd = 2)
       arrows(x0 = OE.cols[,1], y0 = OE.cols[,2], y1 = OE.cols[, 2] + OE.cols[, 3], angle = 90,length = 0.05, lwd = 2)
})
box(lwd=2)

# Construct NM plot
points(NM.cols[,1], NM.cols[,2], type = "b", pch = 22, bg = "white", col = "black",
       cex = 2, cex.lab = 1.5, lwd = 2,
       panel.first={
         arrows(x0 = NM.cols[,1], y0 = NM.cols[, 2], y1 = NM.cols[, 2] - NM.cols[, 3], 
                angle = 90, length = 0.05, lwd = 2)
        arrows(x0 = NM.cols[,1], y0 = NM.cols[, 2], y1 = NM.cols[, 2] + NM.cols[, 3], 
               angle = 90, length = 0.05, lwd = 2)
        abline(v = c(23, 46, 67), col = "black", lty = 3, lwd = 2)
         })

# Add ticks and tick labels
axis(side = 2, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   labels = c("0.000", "-0.005", "-0.010", "-0.015"), 
        at = c(0, -0.005, -0.010, -0.015))

axis(side = 4, labels = F, lwd.ticks = 2, 
   at = c(0, -0.005, -0.010, -0.015))

axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1, mgp = c(3, 2, 0),
    labels = c(0, 25, 50, 75, 100), at = c(0, 25, 50, 75, 100))

axis(side = 3, labels = F, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   at = c(0, 25, 50, 75, 100))

mtext('Water Potential (MPa)', side = 2, outer = TRUE, cex = 1.5, 
      line = -2, adj = 0.6)

mtext('Time (d)', side = 1, outer = TRUE, cex = 1.5, 
      line = -2.5, adj = 0.5)

# Add in-figure caption
text(97, -0.005, labels = "OE", cex = 1.2)
text(97, -0.012, labels = "NM", cex = 1.2)

dev.off()
graphics.off()
```

# Test whether *Pseduomonas* densities are different at end of experiment
```{r}
# Load data
setwd("~/GitHub/BiofilmTrait")
cols.cfu <- read.csv("./data/Biofilm_Columns_CFU.csv", head = TRUE, sep =",")

# Covert CFU of extraction volume to g of soil
vol <- 10 # extraction volume (ml)
soil <- 2.5 # soil extracted (g)
CFUadj <- (cols.cfu$CFU*vol)/soil
CFU <- cbind(cols.cfu, CFUadj)

cols.ttest <- t.test(CFU$CFUadj ~ CFU$Strain)
cols.ttest <- t.test(CFU$CFUadj ~ CFU$Strain, var.equal = TRUE)
stderr <- function(x) sqrt(var(x)/length(x))

OE.cols.CFU <- CFU[which(CFU$Strain=="OE"),]
SEM.OE.cols.CFU <- stderr(OE.cols.CFU$CFUadj)

NM.cols.CFU <- CFU[which(CFU$Strain=="NM"),]
SEM.NM.cols.CFU <- stderr(NM.cols.CFU$CFUadj)
```

## FIG. 6: SPECIES INTERACTIONS

# Read and arrange data
```{r}
# Load data
setwd("~/GitHub/BiofilmTrait")
comp <- read.table("./data/Biofilm_Competition.txt", sep="\t", header=TRUE)
head(comp)

# Calculate growth rates
rate <- (log(comp$tf)-log(comp$t0))/(comp$Hours)
growth <-cbind(comp,rate)
```

# Conduct dummy variables multiple regression (method 1)
```{r}
# subsets KBS0701 growth rate data:
KBS0701 <- growth[which(growth$Strain == "KBS0701"),] 
# dummy var for intercept (D1) and slope (MPa*D1) with 0406:
D1 <- (KBS0701$Culture == "cocult_OE")*1 
# dummy var for intercept (D2) and slope (MPa*D2) on growth with 0407:
D2 <- (KBS0701$Culture!="mono")*1 
# merges KBS0701 data with dummy variables:
KBS0701.2 <- cbind(KBS0701, D1, D2) 
# dummy variables multiple regression:
KBS0701.fit <- lm(rate ~ MPa + D1 + MPa*D1 + D2 + MPa*D2, data = KBS0701.2) 
summary(KBS0701.fit)

# Multiple regression parameters corresponding Psi stars:
tester.int <- KBS0701.fit$coefficients[1]
tester.slp <- KBS0701.fit$coefficients[2]
tester.star <- tester.int/tester.slp*-1

tester.NM.int <- tester.int + KBS0701.fit$coefficients[3]
tester.NM.slp <- tester.slp + KBS0701.fit$coefficients[5]
tester.NM.star <- tester.NM.int/tester.NM.slp*-1

tester.OE.int <- tester.int + KBS0701.fit$coefficients[4]
tester.OE.slp <- tester.slp + KBS0701.fit$coefficients[6]
tester.OE.star <- tester.OE.int/tester.OE.slp*-1

# Dummy variables regresssion: R2 = 0.74, F(5,28) = 16.2, P < 0.0001
# Intercept and Mpa are both significant (0.0003 and <0.0001)
# But no effect of dummary variables on intercepts or slopes (P > 0.22)
```

# Simple linear regression for "global" model (since no effect of strain in multiple regression)
```{r}
global.fit <- lm(KBS0701$rate ~ KBS0701$MPa)
global.int <- global.fit$coefficients[1]
global.slp <- global.fit$coefficients[2]
global.psi <- global.int/global.slp*-1
```

# Plot of growth rates
```{r}
# Strain-specific values
solo <- KBS0701[which(KBS0701$Culture == "mono"),]
with.NM <- KBS0701[which(KBS0701$Culture == "cocult_NM"),]
with.OE <- KBS0701[which(KBS0701$Culture == "cocult_OE"),]

# Main plot
png(filename="~/GitHub/BiofilmTrait/figures/Figure6.png",
    width = 1200, height = 1200, res = 96*2)
par(mar = c(7, 7, 5, 7))
plot(jitter(solo$MPa, factor = 3), solo$rate, xlim = c(0.25, -1.85), 
     ylim = c(-0.1, 0.065),
  pch = 22, bg = "black", col = "black", cex = 2,
  ylab = "", xlab = "", cex.lab = 1.5, cex.axis = 1.2, 
  las = 1, lwd = 2, yaxt = "n", xaxt = "n")
  box(lwd=2)

# Add points
points(jitter(with.NM$MPa, factor = 3), with.NM$rate, pch = 22, 
       bg = "white", col = "black", cex = 2, lwd = 2)
points(jitter(with.OE$MPa, factor = 3), with.OE$rate, pch = 22, bg = "grey", 
       col = "black", cex = 2, lwd = 2)

# Add axis labels
mtext(expression('Growth Rate (d'^-1*')'), side = 2, outer = TRUE, 
      cex = 1.5, line = -3, adj = 0.5)
mtext("Water Potential (MPa)", side = 1, outer = TRUE, cex = 1.5, 
      line = -3, adj = 0.5)

# Add ticks and tick labels
axis(side = 2, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   labels = c("-0.10", "-0.05", "0.00", "0.05"), at = c(-0.1, -0.05, 0.0, 0.05))

axis(side = 4, labels = F, lwd.ticks = 2, 
   at = c(-0.1, -0.05, 0.0, 0.05))

axis(side = 1, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   labels = c("0.0", "-0.5", "-1.0", "-1.5"), at = c(0, -0.5, -1, -1.5))

axis(side = 3, labels = F, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   at = c(0, -0.5, -1, -1.5))

# Add legend
legend(-0.85, 0.07, c("tester", "tester + NM", "tester + OE"), pch = 22, 
       pt.bg = c("black", "white", "grey"), pt.cex = 2, pt.lwd = 2, 
       bty = 'n', y.intersp = 1, lty = c(6,4), lwd = 2.5)

# Add zero-growth line 
abline(h = 0, col = "black", lty = 3, lwd = 2)

clip(0.1, -1.75, -0.1, 0.05)
abline(a = global.int, b = global.slp, col = "black", lwd = 2.5, lty = 6)

# # Add Psi star lines 
segments(x0 = global.psi, y0 = -0.095, x1 = global.psi, y1 = 0, col = "black", 
          lwd = 2.5, lty = 6, xpd = T)

# Add Psi star symbols to plot
mtext(side = 1, line = -1, at = global.psi, bquote(psi~'*'), cex = 1.2)

# Add R2
text(0, 0.06, labels = expression(R^2 == 0.74), cex = 1)

dev.off()
graphics.off()
```



